### 2014-12-01 draft_ck1.tex
another CK edit

%%%% Various options for document class.
%\documentclass[usenatbib, a4paper]{preprint}
\documentclass[preprint, a4paper, 11pt]{aastex}
%%\documentclass[twocolum]{revtex4}
%%\documentclass{report}
%%\documentclass[useAMS,usenatbib, onecolumn]{mn2ejm}

%\usepackage{psfig,morefloats,url}
%use preprint2 for 2 columns paper.

%% declare any packages used
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{color} 
\usepackage{pdfpages}
\usepackage{appendix}
\usepackage{subfigure}
%\usepackage{epsfig}
%\usepackage[dvips]{color}
%\usepackage{aabib}

%\marginparwidth = 25pt
\citestyle{aa}
\addtolength{\topmargin}{-.5in}
%% This command added as margins are wrong in mn2e, it appears. 
%% Not needed for other classes

%% This is the end of the preamble.  Indicate the beginning of the
%% paper itself with \begin{document}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
%% define bibstyle and other definitions
\bibliographystyle{aabib}
%% renew commands
\renewcommand{\labelitemi}{$\bullet$}
\def\py{\textsc{Python} }
\def\tar{\textsc{Tardis} }
\def\cld{\textsc{Cloudy} }
\def\civ{C~\textsc{iv} }
\def\araa{ARAA}
\def\nat{Nature}
\def\apjl{ApJ Letters}
\def\aapr{AAPR}
\def\actaa{ACTAA} 
\def\ssr{SSR}
\def\apj{ApJ}
\def\apss{AP\&SS}
\def\pasp{PASP}
\def\aap{A\&A}
\def\mnras{MNRAS}
\def\aj{AJ}
\def\rmxaa{RMXAA}

%% define some control sequences for lines
\def\heiiopt{He~\textsc{ii}~$\lambda4686{\rm \AA}$}
\def\heiiuv{He~\textsc{ii}~$\lambda1640{\rm \AA}$}
\def\heiioptnew{He~\textsc{ii}~$\lambda3202{\rm \AA}$}
\def\la{Ly-$\alpha$}
\def\ha{H$\alpha$}
\def\hb{H$\beta$}
\def\civfull{C~\textsc{iv}~$\lambda1550{\rm \AA}$}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%          TITLE AND AUTHORS
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{The Impact of Accretion Disk Winds on the Optical Spectra of Cataclysmic Variables}
\author{J. H. Matthews, C. Knigge, K. S. Long, S. A. Sim \& N. Higginbottom}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%          ABSTRACT
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
Many high-state non-magnetic cataclysmic variables (CVs) exhibit
blue-shifted absorption or P-Cygni profiles in ultraviolet
(UV) resonance lines, such as C~{\sc iv}~1550\AA. These features
imply the existence of powerful accretion disk winds in
CVs, and considerable effort has gone into modelling UV line formation
in such outflows. By contrast, much less attention has been paid to
the potential impact of these disk winds on the {\em optical} spectra
of CVs. Here, we use an improved version of our Monte Carlo ionization
and radiative transfer code to investigate whether disk wind models that
produce realistic UV line profiles are also likely to generate
observationally significant recombination line and continuum emission
in the optical waveband. In particular, we test whether outflows may 
be responsible for the single-peaked emission line profiles often seen
in high-state CVs and for the weakness of the Balmer absorption edge
(relative to simple models of optically thick accretion disks).
We find that a standard disk wind model that is successful in
reproducing the UV spectra of CVs does leave a noticeable imprint on
the optical spectrum as well, particularly for systems viewed at high
inclination. The strongest optical wind-formed recombinations lines are \ha\ and
\heiiopt\ , although the outflow also produces emission in the Ly-$\alpha$
and \heiiuv\ UV lines. However, all of the optical lines produced by
this default wind model are double-peaked, and the effect of the wind
on the Balmer jump is modest. We carry out a limited exploration of
the disk wind parameter space, focusing on those parameters that affect the
density near the base of the wind. This shows that higher-density
outflows can produce both single-peaked emission lines and a
recombination continuum sufficient to fill in the Balmer
jump, although such models also tend to predict stronger-than-observed 
collisionally excited UV emission lines. Our main conclusion is that
disk wind emission is likely to have a significant impact on the
optical spectra of high-state CVs. 
\end{abstract}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%          INTRODUCTION
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction} 
\label{sec:intro}

Cataclysmic variables (CVs) are systems in which a white dwarf
accretes matter from a donor star via Roche-lobe overflow. In
non-magnetic systems this accretion is mediated by a Keplerian disk
around the white dwarf (WD). Nova-like variables (NLs) are a subclass
of CVs in which the accretion disk is always in a relatively
high accretion rate state ($\dot{M} \sim
10^{-8}$~M$_{\odot}$~yr$^{-1}$).  This makes NLs an excellent
laboratory for studying the properties of steady-state accretion
disks.

It has been known for a long time that winds emanating from the
accretion disk are important in shaping the ultraviolet (UV) spectra
of high-state CVs (Heap 1978). The most spectacular evidence for such
outflows are the P-Cygni-like profiles seen in UV resonance lines such as
\civfull\ (see e.g. Cordova \& Mason
1982\nocite{cordova1982}). Considerable effort has been spent over the
years on understanding and modelling these UV features (e.g. Drew \&
Verbunt 1985; Mauche \& Raymond
1987; Drew 1987; Shlosman \& Vitello 1993; [hereafter
SV93]\nocite{SV93}; Knigge, Woods \& Drew 1995 [hereafter
KWD95]\nocite{KWD95}; Knigge \& Drew 1997; Knigge et al. 1997; Long \&
Knigge 2002 [hereafter LK02]\nocite{LK02}, Noebauer et al. 2010;
Puebla et al. 2011). The basic picture emerging from these efforts is
of a slowly accelerating, moderately collimated bipolar
outflow that carries away $\simeq 1\% - 10\%$ of the accreting
material. State-of-art-simulations of line formation in this type
of disk wind can produce UV line profiles that are remarkably similar
to observations.

Much less is known about the effect of these outflows on the optical
spectra of high-state CVs. These spectra are typically characterized
by H and He emissions superposed on a blue continuum. In many
cases, and particularly in the SW~Sex subclass of NLs
\citep{HSK86,DR95}, these lines are single-peaked. This is contrary to
theoretical expectations for lines formed in accretion disks, which
are predicted to be double-peaked (REF to Smak and Marsh \&
Horne). {\em Low-state} CVs (dwarf novae in quiescence) do, in fact,
exhibit such double-peaked lines (REF).

Murray \& Chiang (1996) have shown that the presence of disk winds may
offer a natural explanation for the single-peaked optical emission lines in
high-state CVs, since they can strongly affect the radiative transfer
of line photons. Strong support for a significant wind contribution to the
optical emission lines comes from observations of eclipsing
systems. There, the single-peaked lines are often only weakly
eclipsed, and a significant fraction of the line flux remains visible
even near mid-eclipse (REF). This points to line formation in a spatially
extended region, such as a disk wind (see figure~\ref{novalikes}).
%%(Honeycutt, Schlegel, \& Kaitchuck 1986; Dhillon \& Rutten 1995). 
Further evidence for a wind contribution to the optical lines comes
from isolated observations of P-Cygni-like line profiles even in optical
lines, such as \ha\ and He \textsc{i} $\lambda5876$ \citep{RN98}.

Could disk winds also have an impact on the UV/optical {\em continuum}
of high-state CVs? This continuum is usually thought to be dominated
by the accretion disk and modelled by splitting the disk into
a set of concentric, optically thick, non-interacting annuli following
the standard $T_{eff}(R) \propto R^{-3/4}$ radial temperature
distribution (Shakura \& Sunyaev REF). In such
models, each annulus is taken to emit either as a blackbody or,
perhaps more realistically, as a stellar/disk atmosphere model
(REF). In the latter case, the local surface gravity, $\log{g}(R)$, is
assumed to be set solely by the accreting WD, since self-gravity is
negligible in CV disks.

Attempts to fit the observed spectral energy distributions (SEDs) of
high-state CVs with such models have met with mixed success. In
particular, the SEDs predicted by most stellar/disk atmosphere models 
are too blue in the UV (REFs) and exhibit
stronger-than-observed Balmer jumps in absorption (REFs). One possible
explanation for these problems is that these models fail to capture
all of the relevant physics. Indeed, it has been argued that a
self-consistent treatment can 
produce better agreement with observational data (e.g. Shaviv et
al. 1991;  but also more recent work by Idan et al. 2010). 
\nocite{idanshaviv2010} \nocite{shaviv1991}. However, an alternative 
explanation, suggested by \cite{KLWB98}, is that recombination continuum
emission from the base of the disk wind might fill in the disk's
Balmer absorption edge and flatten the UV spectrum.

\begin{figure}	%fullpage
\centering
\includegraphics[width=0.8\textwidth]{figures/fig1.eps}
\caption{Optical spectra of three nova-like variables: 
RW Sex (Top; Beuermann et al. 1992),
IX Vel (Middle; Beuermann \& Thomas 1990)
and RW Tri in eclipse (Bottom; Groot et al. 2004).
The data have been digitized from the respective publications and a few of the more
prominent emission and absorption lines are marked. 
These systems have inclinations of $30^\circ$, $60^\circ$ and $70^\circ$ 
respectively.
The trend of increasing Balmer line emission with inclination can be seen.
In RW Tri strong single-peaked emission in the Balmer lines is seen even
in eclipse, indicating that the lines may be formed in a spatially
extensive disk wind, and there is even a suggestion 
of a (potentially wind-formed) recombination continuum in the eclipsed
spectrum.}
\label{novalikes}
\end{figure}

Here, we carry out Monte Carlo radiative transfer simulations in
order to assess the likely impact of accretion disk winds on the
optical spectra of high-state CVs. More specifically, our goal is to
test whether disk winds of the type developed to account for the UV
resonance lines would also naturally produce significant amounts of  
optical line and/or continuum emission. In order to achieve this, we
have implemented the `macro-atom' approach developed by Lucy
(2002, 2003) into the Monte Carlo ionization and radiative transfer
code described by LK02 (a process initiated by SDL05). With this
upgrade, the code is able to deal correctly with processes involving
excited levels, such as the recombination emission produced by CV
winds.

The remainder of this paper is organized as follows. In Section~2 we
briefly describe the code and and the newly implemented macro-atom
approach. In Section~3, we describe the kinematics and geometry of our
disk wind model. 
In Section~4, we present spectra simulated from the benchmark model
employed by LK02, and in Section~5, we carry out a limited sensitivity
study to explore how the predicted spectra change as a function of
some key parameters in our model. In Section~6, we summarize our
findings.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%          THE CODE, RADIATIVE TRANSFER
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{{\sc python}: A Monte Carlo Ionization and Radiative Transfer Code}

\py is a Monte Carlo (MC) ionization and radiative transfer code which
uses the Sobolev approximation to treat line transfer. It has already
been described extensively in LK02, SDL05 and H13, so here we provide
only a brief summary of its operation, focusing particularly on new
aspects of our implementation of macro-atoms into the code.

\subsection{Basics}

\py operates in two distinct stages. First, the ionization state,
level populations and temperature structure are calculated. This is
done iteratively, by 
propagating several populations of MC energy quanta (`photons')
through a model wind. The geometric and kinemetic properties of the
outflow are specified on a pre-defined spatial grid. In each of these
iterations (``ionization cycles''), the code records estimators that 
characterize the radiation field in each grid cell. At the end 
of each ionization cycles, a new electron temperature is calculated
that more closely balances heating and cooling in the 
plasma. The radiative estimators and updated electron
temperature are then used to revise the ionization state of the wind,
and a new ionization cycle is started. The process is repeated until
heating and cooling are balanced throughout the wind.

This converged model is used as the basis for the second set of
iterations (``spectral cycles''). In these, the emergent spectrum over
the desired spectral range is synthesized by tracking populations of
energy packets through the wind and computing the emergent spectra at
a number of user-specified viewing angles.

\py is designed to operate in a number of different
regimes, both in terms of the scale of the system, and in terms of the
characteristics of the underlying radiation field.
It was originally developed by LK02 in order to model the UV spectra
of CVs with a simple biconical disk wind model. Sim et al. (2005;
hereafter SDL05)\nocite{simmacro2005} used the code to model Brackett
and Pfund line profiles of H in young-stellar objects (YSOs). As part
of this effort, they implemented a `macro-atom' mode (see below) in
order to correctly treat hydrogen recombination lines with
\py. Finally, Higginbottom et al.\nocite{higginbottom2013} (2013;
hereafter H13) used \py to model broad absorption line (BAL) QSOs. For
this application, an improved treatment of ionization was implemented,
so that the code is now capable of dealing with arbritrary
photo-ionizing SEDs, including non-thermal and multi-component ones.

\subsection{Ionization and Excitation: ``Simple Atoms''}

Prior to SDL05, the relative ionization fractions for all atomic
species were estimated via the modified Saha equation (Mazzali \&
Lucy 1993)  
\begin{equation}
\frac{n_{j+1} n_e}{n_j} = W [\xi + W(1-\xi)]
\left(\frac{T_e}{T_R}\right)^{1/2}
\left(\frac{n_{j+1}n_e}{n_j}\right)^*_{T_R}, \label{ionization}
\end{equation}
Here, the `starred' term on the right represents Saha
abundances computed at temperature $T_R$, but
using actual partition functions obtained 
from the dilute blackbody approximation.
This is in accordance with the notation of Mihalas (1982).  
%%JMTASK: in accordance with the notation of Mihalas (1982).  - check this!
$W$ is an effective dilution factor, $\xi$ is the
fraction of recombinations going directly to the ground state, and
$T_R$ and $T_e$ are the radiation and electron temperature
respectively. This simple ionization scheme produces reasonable
results when the photoionizing SED can be approximated by a dilute
blackbody. This is the case for high-state CVs. (As noted above, an
improved, but more complex treatment of ionization that is appropriate
for more complex SEDs is described in H13.)

Similarly, the relative excitation fractions within each ionization
stage of a given species were estimated via a modified Boltzmann
equation,
\begin{equation}
\frac{n_{jk}}{n_j} = W g_k \exp{-E_k/kT_R} / z_j(T_R),
\end{equation}
where $n_{jk}$ is the population of level $k$ in ionic stage $j$,
$E_k$ is the energy difference between level $k$ and the ground state,
and $z_j(T_R)$ is the partition function of ionic stage $j$.

Finally, \py modelled all bound-bound processes as transitions
within a simple two-level atom \cite[see e.g.][]{mihalas}. This
approach works reasonable well for resonance  
lines, such as \civfull, in which the lower level is the ground state.  
However, it is a poor approximation for many other
transitions, particularly those where the upper level is 
is primarily populated from above. Thus an improved method for
estimating excited level populations and simulating line transfer is
needed in order to model recombination lines and continua.

\subsection{Ionization and Excitation: Macro-Atoms}

Lucy (2002, 2003\nocite{lucy2002, lucy2003}; hereafter L02, L03) 
has shown that it is possible to calculate the emissivity of a gas in
statistical equilibrium accurately by quantising matter into
`macro-atoms', and radiant and kinetic energy into indivisible energy
packets (r- and k- packets, respectively). His macro-atom scheme
allows for all possible transition paths from a given level and
provides a full non-local thermodynamic equilibrium (NLTE) solution
for the level populations based on MC estimators. The macro-atom
technique has already been used to model Wolf-Rayet star
winds \citep{sim2004}, AGN disk winds \citep{simlong2008, tatum2012},
supernovae \citep{kasen2006, kerzendorfsim} and YSOs (SDL05). A full 
description of the scheme can be found in L02 and L03.

Briefly, macro-atom NLTE level populations and ionization fractions
are calculated by solving the statistical equilbrium equations between
each pair of levels. For example, the bound-bound de-excitation rate,
${\cal R}_{ul}$, in an ion is given by  
\begin{equation}
{\cal R}_{ul} = \beta_{lu} A_{ul} n_u + \beta_{lu} B_{ul} n_u J_{lu} +
C_{ul} n_u n_e,\label{eq:nlte_rul},
\end{equation}
where $u$ and $l$ denote the upper and lower levels, $C$ represents the
collisional rate coefficients and $A$ and $B$ represent the Einstein
coefficients. The quantity $\beta_{lu}$ is the probability that a
given line photon will escape the Sobolev region. Finally, $J$ is the
MC estimator for the mean intensity in the line, as recorded during
the photon propagation. The corresponding excitation rate is then 
\begin{equation}
{\cal R}_{lu} = \beta_{lu} B_{lu} n_l J_{lu} + C_{lu} n_l n_e,
\end{equation}
where we implicitly include a correction for stimulated emission in
our definition of the estimate for $J_{ul}$.

In our implementation of the macro-atom approach, we also explicitly
take into account the photoionization and collisional ionization rates
between a lower level, $l$, and the continuum (or, in the case of ions
with more than one bound electron, the ground state of the upper ion),
$\kappa$,
\begin{equation}
{\cal R}_{l \kappa}= n_l \int_{\nu_0}^{\infty} \frac{ 4 \pi J_{\nu}
  \sigma_{\nu}}{h \nu} d\nu + C_{l \kappa} n_l n_e.
\end{equation}

Here, $\sigma_{\nu}$ is the photoionization cross section, and $J_{\nu}$
is the mean intensity. The corresponding recombination rate is given
by 
\begin{equation}
{\cal R}_{\kappa l} = \alpha_{\kappa l} n_{\kappa} n_e + C_{\kappa l}
n_\kappa n_e, \\
\end{equation}
where $\alpha_{\kappa l}$ is the radiative recombination coefficient
to level $l$. This treatment means that radiative and collisional
rates to and from all levels are considered when calculating both the
ionization state and the level populations, although we neglect 
ionization directly to excited levels of the upper ion. The
\cite{vanregemorter} approximation is used for collisional
transitions, and thus we assume that collisions between radiatively
forbidden transitions are not important in determining level
populations. This is a reasonable approximation in the regimes
considered here.

\subsection{Ionization and Excitation: A Hybrid Approach}

SDL05 implemented a macro-atom treatment of Hydrogen in \py and used
this to predict the observable properties of a pure Hydrogen wind
model for YSOs. Our goal here is to simultaneously model the optical
and ultraviolet spectra of high-state CVs. Since the optical spectra
are dominated by H and He recombination lines, both of these species
need to be treated as macro-atoms. The UV spectra, on the other hand,
are dominated by resonance lines associated with metals. This means we
need to include these species in our models, but they can be treated 
with our (much faster) simple-atom approach. We have therefore
implemented a hybrid ionization and excitation scheme into \py. Any
atomic species can now be treated either in our simple-atom
approximation or with the full macro-atom machinery. For our CV
models, we treat H and He as macro-atoms and all metals as
simple-atoms.

\subsection{Atomic Data}

We generally use the same atomic data as H13, which is an updated
version of that described in LK02. In addition, we follow SDL05 in
treating Hydrogen as a 20-level atom, where each level is defined by
the principal quantum number, $n$. For the macro-atom treatment of
Helium,  we have added additional level and line information required 
from \textsc{Topbase} \citep{topbase2005}.  He~\textsc{II} is treated
in much the same way as H, but with 10 levels. He~\textsc{I} has
larger energy differentials between different l-subshells and triplet
and singlet states. Thus, we still include levels up to $n=10$ but
explicitly treat the $l$ and $s$ sub-orbitals as distinct levels
instead of assuming they are perfectly `l-mixed'. This allows us
to model the singlet and  triplet He~{\sc i} lines that are ubiquitous
in the optical spectra of CVs \citep[e.g.][]{dhillon1996}.

\subsection{Code Validation and Testing}

\py has been tested against a number of radiative transfer and
photoionization codes. LK02 and H13 conducted comparisons of 
ionization balance  with \cld \citep{cloudy2013}, demonstrating
excellent agreement. We have also carried out comparisons
of ionization and spectral synthesis with the supernova code
\textsc{Tardis.} \tar is described by 
\cite{kerzendorfsim}, and the spectral comparisons can be found
therein. For the effort reported here, we have additionally carried
out tests of the macro-atom scheme in \py. Figure~\ref{tests} shows
two of these tests. In the top panel, we compare the Balmer series 
emissivities as predicted by \py in the l-mixed Case~B limit against the
analytical calculations by \cite{seaton1959}. In the bottom panel, we
compare \py and \tar predictions of  He \textsc{i} level populations
for a particular test case. Agreement is excellent for both H and He.

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/fig_caseb_tardis.eps}
\caption{
{\sl Top Panel:} Balmer decrements calculated with \py compared with analytic calculations
by Seaton (1959), showing good agreement when calculating level emissivities in Hydrogen
in the Case B limit.
(see Osterbrock 1989 for a discussion of this commonly used approximation).
{\sl Bottom Panel:}  a comparison of Helium I level populations (the most complex ion we currently 
treat as a macro-atom) between \py~and \tar~models. Considering the 
two codes use different atomic data and \textsc{Tardis,} unlike \textsc{Python,} currently has 
a complete treatment of collisions between radiatively forbidden transitions, the factor of 
$<2$ agreement is encouraging. 
{\bf CK: WHAT ARE THE PARTICULAR SETS OF PARAMETERS BEING ADOPTED HERE?}
}
\label{tests}
\end{figure}
\nocite{osterbrock}
\nocite{seaton1959}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%          THE MODEL
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Describing the System and its Outflow}

\py includes several different kinematic models of accretion disk
winds, as well as different options for describing the physical and
radiative properties of the wind-driving system under
consideration. Most of these features have already been discussed in
LK02 and H13, so below we only briefly recount the key aspects of the
particular system and wind model used in the present study.

\subsection{Wind Geometry and Kinematics}

We adopt the kinematic disk wind model developed by Schlosman \&
Vitello (1993; hereafter SV93). A schematic of this model is shown in
figure~\ref{cartoon}. In this parametrization, a smooth, biconical
disk wind emanates from the accretion disk between radii $r_{min}$ and 
$r_{max}$. The covering fraction of the outflow is also controlled by the
inner and outer opening angles of the wind, $\theta_{min}$ and
$\theta_{max}$, and the launch angle of the other streamlines is given
by 
\begin{equation}
\theta(r_0) = \theta_{min} + (\theta_{max} - \theta_{min}) \left(\frac{r_0 - r_{min}}{r_{max} - r_{min}} \right)^{\gamma},
\label{theta}
\end{equation}
where $r_0$ is the launch radius of the streamline.

The poloidal (non-rotational) velocity field of the wind, $v_l$, is given by
\begin{equation}
v_l=v_0+\left[v_{\infty}(r_0)-v_0\right]\frac{\left(l/R_v\right)^{\alpha}}{\left(l/R_v\right)^{\alpha}+1},
\label{v_law}
\end{equation}
where $l$ is the poloidal distance along a particular wind
streamline. The terminal velocity along a streamline, $v_{\infty}$, is
set to a fixed multiple of the escape velocity at the launch
point. The launch velocity from the disk surface, $v_0$, is assumed to
be constant (set to $6$~km~s$^{-1}$). Once the wind is launched, it
accelerates, reaching half of its terminal velocity at $l = R_v$. The
velocity law exponent $\alpha$ controls how quickly the wind
accelerates. Larger values of $\alpha$ cause the main region of 
acceleration to occur close to $R_v$, whereas smaller values
correspond to fast acceleration close to the disk (see
figure~\ref{acc_law}).

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/fig2_cartoon.eps}
\caption{Cartoon illustrating the geometry and kinematics of the benchmark CV wind model.}
\label{cartoon}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{figures/acc_law.eps}
\caption{
The adopted poloidal velocity law for various values of the
acceleration exponent, $\alpha$.
}
\label{acc_law}
\end{figure}

The density at position $(r,z)$ in the wind, $\rho(r,z)$, is
calculated from the mass continuity equation, yielding
\begin{equation}
\rho(r,z) = \frac{r}{r_0} \frac{dr}{dr_0} \frac{\phi(r_0)}{v_z(r,z)}.
\label{density}
\end{equation}
Here, $\phi(r_O)$ is the mass-loss rate per unit area at $r_0$, 
and $v_z$ is the vertical velocity component. We take $\phi(r_0)$ to be
uniform across the wind-launching area of the disk and normalize it by
matching its integral to the user-specified total mass-loss rate,
$\dot{M}_{wind}$. 
{\bf I'm still not sure I understand the formula
  SV93 use to define $\phi$, and hence that this description is
  actually strictly correct.... Not sure what to do about that
  though...}

\subsection{Sources and Sinks of Radiation}

The net photon sources in our CV model are the accretion disk, the
WD and, in principle, a boundary layer with user-defined temperature
and luminosity. All of these radiating bodies are taken to be
optically thick, and photons striking them are assumed to be destroyed
instantaneously. The secondary star is not included as a radiation
source, but is included as an occulting body. This allows us to model
eclipses. Finally, emission from the wind itself is also accounted for, but
note that we assume the outflow to be in radiative equilibrium. Thus all
of the heating of the wind, as well as its emission, is ultimately
powered by the radiation field of the net photon sources in the
simulation. In the following sections, we will describe our treatment
of these system components in slightly more detail.

\subsubsection{The Accretion Disk}

\py has has some flexibility when treating the accretion 
disk as a source of photons. The disk is broken down into annuli 
such that each annulus contributes an equal amount to the bolometric
luminosity. We take the disk to be geometrically thin, but optically
thick and thus adopt the temperature profile of a standard
\cite{shakurasunyaev1973} $\alpha$-disk. An annulus can then
be treated either as a blackbody with the corresponding effective
temperature or as a stellar atmosphere model with the appropriate
surface gravity and effective temperature. Here, we use blackbodies 
during the ionization cycles and to compute our MC
estimators. However, during the spectral synthesis stage of the 
simulation we use stellar atmosphere models. This produces more
realistic model spectra and allows us to test if recombination
emission from the wind base can fill in the Balmer jump, which is
always in absorption in these models. Our synthetic stellar atmosphere
spectra are calculated with
\textsc{Synspec}\footnote{http://nova.astro.umd.edu/Synspec43/synspec.html}
from either Kurucz \citep{kurucz1991} atmospheres (for $T_{eff} \leq
50,000$~K) or from \textsc{TLUSTY} models \citep{tlusty} (for $T_{eff} > 50,000$~K).

\subsubsection{The White Dwarf}

The WD at the center of the disk is always present as a spherical occulting
body with radius $R_{WD}$ in \py CV models, but it can also be included
as a source of radiation. In the models presented here, we treat the
WD as a blackbody radiator with temperature $T_{WD}$ and luminosity
$L_{WD} = 4\pi R_{WD}^2 \sigma T_{WD}^4$.

\subsubsection{The Boundary Layer}

It is possible to include radiation from a boundary layer (BL) between
the disk and the WD. In \py, the BL is described as
a blackbody with a user-specified effective temperature and
luminosity. In the models presented here, we have followed LK02 in setting
the BL luminosity to zero. However, we have confirmed that the addition of a
BL with $L_{BL} = 0.5 L_{acc}$ and temperatures in the range $80{\rm
kK} \leq T_{BL} \leq 200{\rm kK}$ would not change any of our main
conclusions. 
{\bf CK: Is this true? Do we want to say anything else, e.g. that we'll
     talk about this in a future paper?}

\subsubsection{The Secondary Star}

The donor star is included in the system as a pure radiation sink, 
i.e. it does not emit photons, but absorbs any photons that strike its
surface. The secondary is assumed to be Roche-lobe filling, so its
shape and size are defined by setting the mass ratio of the system, $q
= M_2/M_{WD}$. The inclusion of the donor star as an occulting body
allows us to model eclipses of the disk and the wind. For this
purpose, we define orbital phase such that $\Phi_{orb} = 0$ as the
inferior conjunction of the secondary (i.e. mid-eclipse for $i \simeq
90^o$).

\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{figures/sed_figure.eps}
\caption{
The spectral energy distribution (SED) of the blackbody accretion
disk used in the ionization calculation (black line), 
with two boundary layer blackbody spectra also plotted.
The positions of the Hydrogen and Helium ionization edges 
are marked with dashed vertical lines.
{\bf CK: this will need to be updated -- we want to show (just) the disk and
  WD spectra...}
}
%%JMTASK--CK: this will need to be updated -- we want to show (just) the disk and WD spectra...
\label{sed}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%          RESULTS
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}
\centering
\begin{tabular}{p{3cm}p{4cm}}
Model A \\
\hline Free Parameters 	&	 Value \\ 
\hline \hline 
$M_{WD}$ 	 &	 $0.8 M_{\odot}$ \\ 
$R_{WD}$ 	 &	 $7\times10^{8}$cm\\ 
$T_{WD}$ 	 &	 $40,000$K\\
$M_{2}$ 	 &	 $???? M_{\odot}$ \\ 
$\dot{M}_{acc}$ 	 &	 $10^{-8}~M_{\odot}yr^{-1}$\\ 
$\dot{M}_{wind}$  &	$10^{-9}~M_{\odot}yr^{-1}$å\\ 
$r_{min}$ 	&	 $4 R_{WD}$\\ 
$r_{max}$ 	&	 $12 R_{WD}$ \\ 
$\theta_{min}$ 	&	 $20.0^{\circ}$ \\ 
$\theta_{max}$ 	&	 $65.0^{\circ}$ \\ 
$\gamma$ 	&	 $1$ \\ 
$v_{\infty}$ 	&	 $3v_{esc}$ \\ 
%%$R_v$ 	        &	 $7\times10^{10}$cm \\ 
$R_v$ 	        &	 $100 R_{WD}$ \\ 
$\alpha$ 	&	 $1.5$ \\
\end{tabular}
\centering
\caption{
Parameters used for the geometry and kinematics of the benchmark CV model.}
\label{wind_param}
\end{table}

\section{A Benchmark CV Disk Wind Model}

Our main goal is to test whether the type of disk wind model that has
been successful in explaining the UV spectra of CVs could also have 
significant impact on the optical continuum and emission line spectra
of these systems. In order to set a benchmark, we therefore begin by
investigating one of the fiducial CV wind models that was used by SV93
and LK02 to simulate the UV spectrum of a typical high-state
system. The specific parameters for this model are listed in
Table~1. A key point is that the wind mass-loss rate in this model is
set to 10$\%$ of the accretion rate through the disk.

\subsection{Physical Structure and Ionization State}

Figure~\ref{wind} shows the physical and ionization structure 
of the benchmark disk wind model. There is an obvious drop-off in density
and temperature with distance away from the disk, so any line
formation process that scales as $n_e^2$ -- i.e. recombination and
collisionally excited emission -- should be expected to operate
primarily in the dense base of the outflow. Moreover, a comparison of
the rotational and poloidal velocity fields shows that rotation
dominates in the near-disk regime, while outflow dominates further out
in the wind.

The ionization equation used in the ``simple atom'' approach used in
LK02 (see Section~\ref{BLA}) should be a reasonable approximation to
the photoionization equilibrium in the benchmark wind model. Even
though the macro-atom treatment of H and He does affect the 
computation of the overall ionization equilibrium, we would expect the
resulting ionization state of the wind to be quite similar to that
found in LK02. The bottom panels in Figure~\ref{BL} confirm that this
is the case. In particular, Helium is fully ionized
throughout most of the outflow, except for a small region near the
base of the wind, which is shielded from the photons produced by the
hot inner disk. In line with the results shown in LK02, we also find
that C\textsc{iv} is the dominant Carbon ion throughout the wind,
resulting in a substantial absorbing column across a large range of
velocities. As we shall see, this produces the broad, deep and
blue-shifted C\textsc{iv}~$\lambda 1550{\rm \AA}$ absorption line that
is usually the most prominent wind-formed feature in the UV spectra of
low-inclination nova-like CVs.

\subsection{Synthetic Spectra}

We begin by verifying that the benchmark model still produces UV
spectra that resemble those observed in CVs. We do expect this to be
the case, since the ionization state of the wind has not changed
significantly from that computed in LK02 (see Section~\ref{BLA}). The
left column of panels in Figure~\ref{spec} shows that this expectation
is met: all of the strong metal resonance
lines -- notably N\textsc{v}~$\lambda 1240{\rm \AA}$,
Si\textsc{iv}~$\lambda 1400{\rm \AA}$ and C\textsc{iv}~$\lambda
1550{\rm \AA}$ -- are present and exhibit clear P-Cygni profiles
at intermediate inclinations. In addition, however, we now also find
that the wind produces significant Lyman-$\alpha$ and
He~\textsc{ii}~$\lambda1640{\rm \AA}$ emission lines.

\begin{figure} %fullpage
\includegraphics[width=\textwidth]{figures/fig5_uv_opt.eps}
\caption{
UV (left) and optical (right) synthetic spectra for model A, our benchmark model,
computed at sightlines of 10, 27.5, 45, 62.5 and 80 degrees.	
The inset plots show zoomed-in line profiles for 
\heiiuv\ and \ha. Double-peaked line emission can be seen in 
\heiiuv, \heiiopt, \ha\ and some He I lines, but the 
line emission is not always sufficient to overcome the absorption
cores from the stellar atmosphere models. The model
also produces a prominent \heiioptnew\ line at high inclinations,
}
\label{spec}
\end{figure} %fullpage\

Figure~\ref{spec} shows the corresponding optical spectra produced for
the benchmark model, and these do exhibit some emission lines
associated with H and He. Since the emissivity of these recombination 
features scales as $n_e^2$, they are formed almost entirely in the 
dense base of the wind, just above the accretion disk. Here, the
velocity field of the wind is still dominated by rotation, rather than
outflow, which accounts for the double-peaked shape of the lines. In
principle, lines formed in this region can still be single peaked,
since the existence of a poloidal velocity {\em gradient} changes the
local escape probabilities (Murray \& Chiang 1996). However, as
demonstrated explicitly in Section~\ref{BLA}, the line opacity in our
benchmark model is not high enough for this radiative transfer effect
to dominate the line shapes.

We also see a general trend from absorption lines to emission lines 
with increasing inclination, as one might expect from our wind
geometry. This trend is consistent with observations, as can be seen
in Figure~1.

The Balmer jump is in absorption at all inclinations for our benchmark
model. This is due to the stellar atmospheres we have used to
model the disk spectrum; it is not a result of photoabsorption in the
wind. In fact, the wind spectrum exhibits the Balmer jump in {\em
emission}, but this is not strong enough to overcome the intrinsic
absorption edge in the disk spectrum. This is illustrated in
Figure~\ref{cont}, which shows the angle-integrated spectrum of the system,
i.e. the spectrum formed by all escaping photons, separated into the
disk and wind contributions. Even though the wind-formed Balmer
recombination continuum does not completely fill in the Balmer
absorption edge in this model, it does already contribute
significantly to the total spectrum. This suggests that modest changes 
to the outflow kinematics might boost the wind continuum and produce
emergent spectra with weak or absent Balmer absorption edges.

\begin{figure} 
\includegraphics[width=0.45\textwidth]{figures/fig6_opt_cont.eps}
\caption{Synthetic optical spectra computed for 
sightlines of 22.5, 25, 62.5 and 80 degrees. In these plots
the flux is divided by a polynomial fit to the 
underlying continuum redward of the Balmer edge, so that 
line-to-continuum ratios and the true depth of the
Balmer jump can be shown.}
\label{spec_continuum}
\end{figure}

\begin{figure} 
\includegraphics[width=0.5\textwidth]{figures/fig7_escaping.eps}
\caption{Total packet-binned spectra across all viewing angles. 
The thick black line shows the total 
integrated escaping spectrum, while green and red lines show the contributions from photons that originate
in the disk and wind respectively. Recombination continuum emission blueward of the Balmer 
edge is already prominent relative to other wind continuum processes, but is not sufficient
to fill in the Balmer jump in this specific model}
\label{cont}
\end{figure}

\begin{figure} %fullpage
\includegraphics[width=\textwidth]{figures/fig5.eps}
\caption{
The physical properties of the wind- note the logarithmic scale. 
Near the disk plane the wind is dense, with low poloidal velocities.
As the wind accelerates it becomes less dense
and more highly ionized. The dominant Helium ion
is almost always He III, apart from in a small
portion of the wind at the base, which is partially shielded
from the inner disk.
}
\label{wind}
\end{figure} %fullpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%          DISCUSSION
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{The Dependence of Wind Features on Outflow Parameters}

The results of Section~\ref{BLA} show that wind models that reproduce
the UV spectral features of CVs are likely to leave a clear imprint on
the optical spectrum. It is therefore important to understand the
sensitivity of optical (and UV) wind-formed features to the basic
properties of the outflow. In particular, there are two key questions
we would like to address: (i) under what conditions can disk winds
produce {\em single-peaked} Balmer emission lines (Murray \& Chiang
1996); (ii) can the wind-formed recombination continuum completely
fill in the disk's Balmer absorption edge for reasonable outflow
parameters?

The amount of recombination and collisionally excited emission
produced by a given ion, $i$, in a plasma of volume $V$ depends
primarily on the emission measure, $EM$, defined as 
\begin{equation}
EM=\int^V_0 n_e n_i \,dV,
\end{equation}
where $n_i$ is the number density of the ion in question.    
The emission reaching an observer at a given frequency also depends on
the Sobolev optical depth it experiences, which depends on both the
density and the velocity gradient projected along the line of
sight. This can also significantly affect line profile shapes.

The simplest way to simultaneously affect the density in the wind as
well as the velocity gradients is by modifying the poloidal velocity
law. We have therefore conducted a exploratory sensitivity study in
which we focus on just two kinematic variables (Section~\ref{BLA}):
\begin{itemize}
 	\item the acceleration length, $R_v$, which controls the
        distance over which the wind accelerates to $\frac{1}{2}v_{\infty}$;
 	\item the acceleration exponent, $\alpha$, which controls the rate 
 	at which the poloidal velocity changes near $R_v$.
\end{itemize} 
The general behaviour we might expect is that outflows with denser
regions near the wind base -- i.e. winds with larger $R_{v}$ and/or
larger $\alpha$ -- will produce stronger emission signatures. However,
this behaviour may be moderated by the effect of the increasing
optical depth through this region.

\subsection{Line Profiles: Shapes and Intensities}

Figure~\ref{halpha} shows how the H$\alpha$ and He~{\sc ii}~$\lambda
4686$ lines change with the kinematics of the wind for an inclination
of $80^\circ$. The width and strength of He~{\sc ii} clearly depend on
both $R_v$ and $\alpha$, but the overall line profile shape is always
double-peaked. This contrasts with the behaviour of H$\alpha$, which
becomes single-peaked in our densest models. The transition from
double-peaked to single-peaked occurs when the line becomes optically
thick. This reproduces the behaviour predicted by \cite{MC96}, who
showed that the poloidal velocity gradient in the wind can modify the
shape of optically thick emission lines even in the near-disk regime,
where the poloidal velocity itself is much smaller than the rotational
velocity. This has the effect of suppressing the line wings and
can therefore lead to the production of a single-peaked line profile
shape.

In order to produce single-peaked line emission in the majority of
lines at high inclinations, even higher opacity is needed in the
line wings. This difficulty in generating single-peaked line emission
from a rotating disk wind was also encountered by SDL05, who found
that an outflow geometry similar to that adopted here could not 
reproduce the observed single peaked profiles in YSOs. For the
parameters we have tested, the mechanism proposed by \cite{MC96} does
produce single-peaked profiles, but only for most optically thick
emission lines. In order for the same mechanism to affect the majority
of lines, even higher wind densities would be required.

{\bf CK: done up to here -- still need to edit discussion of Balmer
  jump and CIV, and then the ``next-gen'' model''}

\begin{figure} %fullpage
\mbox{
\subfigure{\includegraphics[width=0.5\textwidth]{figures/3by3_grid_alpha.eps}}
\quad
\subfigure{\includegraphics[width=0.5\textwidth]{figures/3by3_grid_beta.eps}}   
}
\caption{
Line profiles for \ha\ (left) and \hb\ (right) with varying kinematic 
properties, computed for an inclination of $80^\circ$.
The benchmark model is shown in the top left panel, and the improved optical
model described in section 6 is highlighted in bold. 
The parameter that has been altered from the benchmark model 
is marked in each column and row. 
Models towards the bottom right
are have more slowly accelerating winds and denser wind bases. }
\label{halpha}
\end{figure} %fullpage

\subsection{The Balmer Jump}

\label{balmerjump}

The left panel of figure~\ref{jump} shows the evolution
of the Balmer jump with our kinematic parameters. 
The `absorption' feature just redward of the Balmer jump
is artificial. It is caused by line blanketing in the stellar atmospheres,
which is not filled in by line emission from the wind
due to the use of a 20-level Hydrogen atom.
The higher density models
produce sufficient recombination continuum emission to fill 
in the Balmer jump. This offers a potential solution to a 
longstanding problem when reconciling theoretical
models of NL spectra with observations, as originally suggested by
\cite{KLWB98}. The fact that this emission is very sensitive to the parameters
that govern the density at the wind base also supports
the idea that a transition region close to the disk plane is necessary
to explain the emission features present \citep{kd1997}.

\begin{figure} %fullpage
\mbox{
\subfigure{\includegraphics[width=0.5\textwidth]{figures/3by3_grid_1640.eps}}
\quad
\subfigure{\includegraphics[width=0.5\textwidth]{figures/3by3_grid_4686.eps}}   
}
%%\includegraphics[width=0.5\textwidth]{figures/test.eps}
\caption{
Line profiles for \heiiuv\ (left) and \heiiopt\  (right) with varying kinematic 
properties, computed for an inclination of $80^\circ$. 
The panels are organised in the same way as figure~\ref{halpha}.
% The panels are ordered in the same way as figure~\ref{halpha}. In general, increasing densities
% lead to a Balmer jump in emission, although as in the case of H$\alpha$ there is a transition
% when the wind becomes a neutral absorbing column.
}
\label{heiifig}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%          Comparison with real spectra
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{An Improved Optical Model }

\begin{table}
\centering
\begin{tabular}{p{3cm}p{4cm}}
Model B \\
\hline Free Parameters 	&	 Value \\ 
\hline \hline 
$M_{WD}$ 	 &	 $0.8 M_{\odot}$ \\ 
$\dot{M}_{acc}$ 	 &	 $10^{-9}~M_{\odot}yr^{-1}$\\ 
$\dot{M}_{wind}$  &	$10^{-8}~M_{\odot}yr^{-1}$\\ 
$r_{min}$ 	&	 $4 R_{WD}$\\ 
$r_{max}$ 	&	 $12 R_{WD}$ \\ 
$\theta_{min}$ 	&	 $20.0^{\circ}$ \\ 
$\theta_{max}$ 	&	 $65.0^{\circ}$ \\ 
$\gamma$ 	&	 $1$ \\ 
$v_{\infty}$ 	&	 $3v_{esc}$ \\ 
$R_v$ 	        &	 $142.9 R_{WD}$ \\ 
$\alpha$ 	&	 $4$ \\
\end{tabular}
\centering
\caption{Wind geometry parameters used in the next-generation CV model.}
\label{modelb}
\end{table}

With modest changes to the kinematics of the benchmark CV model,
it is possible to produce a model
that comes closer to reproducing the observed optical spectrum
of a NLs. The parameters for this model (imaginatively dubbed Model B)
are show in table~\ref{modelb}, and the synthetic spectrum
is shown in figure~\ref{uvoptb}. An in and out of eclipse comparison 
between the synthetic spectrum
at an angle of $80^\circ$ and the high-inclination NL RW Tri 
is also shown in figure~\ref{rwtricomp}. The inclination
for RW Tri has been disputed, with claimed values of $75^\circ$ \citep{groot2004},
$70.5^\circ$ \citep{smak1995}, $80^\circ$ \citep{longmore1981} and 
$82^\circ$\citep{frankking1981}. A lower inclination
would mean that our model would simply require lower densities 
and smaller emission regions to produce the required continuum level
and single peaked emission, so our conclusions from this comparison are
not particular sensitive to the value chosen.
We also emphasize that this is no sense a fit to the data. The model
is purely designed to reproduce the general behaviour of NLs
and give an indication of the likely densities and kinematics 
required in the wind.
%quotes a figure of , whereas  halevinhenden2008

The similarity with the observed spectrum is striking. 
Model B produces strong emission in all the Balmer lines, 
with line to continuum ratios
comparable to those in RW Tri. 
The line to continuum ratio increases during eclipse,
as one expects if the emission is produced in a disk wind, 
thereby mimicking the behaviour of SW Sex stars.
The single-peaked \ha\ line is produced by our models with a 
very similar width to the observed spectrum, although it is perhaps slightly more
`flat-topped' in the synthesized data.

One of the limitations of the model is that, apart 
from in H$\alpha$, the lines are in general double-peaked. 
The lack of high-resolution data in this regard makes detailed analysis difficult,
but it would appear that understanding how single-peaked 
line emission forms and the kinematics and densities required to
modify the line profile is still an open question. 
The major fallback of the next-generation model, however, lies
in the ultraviolet. To get the required emission in the optical lines
and recombination continuum the densities required are very high 
($n_e\sim10^{13}-10^{14}$ in the main emission region).
This causes a significant collisionally excited contribution
to the UV lines, meaning that the red wing of the 
resonance species such as C\textsc{iv} no longer matches the
observed profiles. {\bf ksl - I do not think it follows that because the red wing is strong, that this is due to collisional effects.  I think your wind is now effectively just more extended and this produce more scattered radiation. Scttering goes like n(C), emission goes like ne*n(C).  It would be interesting if we could figure out a way to show graphically where emission and absorption in a wind model are coming from.}

\begin{figure} %fullpage
\includegraphics[width=\textwidth]{figures/fig14_uv_opt.eps}
\caption{
UV (left) and Optical synthetic spectra for model B computed at
sightlines of 10, 27.5, 45, 62.5 and 80 degrees.	
The inset plots show zoomed-in line profiles for 
\heiiuv\ and \ha. The \ha line 
is single peaked, but higher order lines in the Balmer series
are double-peaked, albeit with narrower profiles.
Strong \heiiopt\ emission can be seen, as well a trend
of a deeper Balmer jump with decreasing inclination.
}
\label{uvoptb}
\end{figure} %fullpage

\begin{figure} %fullpage
\includegraphics[width=0.8\textwidth]{figures/fig13_eclipse.eps}
\caption{{\sl Top Panel:} In and out of eclipse spectra of the high
inclination NL RW Tri. {\sl Bottom Panel:} In and out of eclipse synthetic
spectra from Model B.
The artificial `absorption' feature just redward of the Balmer jump
is caused for the reasons described in section 5.2.}
\label{rwtricomp}
\end{figure} %fullpage

% \noindent\rule{16cm}{0.4pt}

% {\bf
% \noindent Notes on Section 6:

% \begin{itemize}
% 	\item

% \noindent Potential Additional Figures:

% \begin{itemize}
% 	\item 	
% }

% \noindent\rule{16cm}{0.4pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%          Conclusions and Future Work
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusions}

By taking an already successful model
for the UV wavelength regime of CVs and treating 
it with an improved line transfer scheme, we have shown
that the wind does produce some recombination lines
and continuum emission blueward of the Balmer edge.
We have confirmed a number of the expected behaviours
by exploring a limited number of parameters governing the
kinematics and geometry of the wind. These include:

\renewcommand{\labelitemi}{$\bullet$}
\begin{itemize}
	\item The `filling-in' of the Balmer jump \citep{KLWB98} by recombination 
	continuum emission from a disk wind
	\item Single peaked line emission in \ha\ produced by 
	the introduction of a poloidal velocity gradient \citep{MC96}
	\item A Balmer decrement consistent with optically thick emission \citep{elitzur1983}.
\end{itemize}
\smallskip

\noindent We also make a number of predictions:

\begin{itemize}
	\item The depth of the Balmer absorption edge decreases with increasing inclination,
or alternatively that the Balmer recombination emission
is enhanced at higher inclinations, {\sl relative to the disk continuum}.
	\item The He~\textsc{ii}~$\lambda3202{\rm \AA}$ line should be
prominent in cases where He~\textsc{ii}~$\lambda1640{\rm \AA}$ \& 
\heiiopt\ are seen. 
	\item We expect Balmer jump in emission during eclipse.
\end{itemize}

\smallskip
Our improved optical model produces all the prominent lines in and out of eclipse, and
exhibits verisimilitude with the optical spectrum of RW Tri.
However, this model does not produce a complete unified CV model. 
The emission lines in the UV are somewhat unrealistic in strength,
because the high densities needed to increase line opacity also produce
a significant collissionally excited contribution..
This is a shortcoming which must be understood if it is true that
disk winds modify line profiles to become single peaked.
% as the 
%densities required produce a large amount of collisionally excited emission
%in UV resonance lines.

Our work demonstrates that, if nothing else,
{\sl disk winds matter} in the optical. At a minimum,
they can act as a spectral filter
for the disk spectrum.
%and the properties of the plasma
%is an important tracer for the disk and boundary layer SEDs. 
At a maximum, they may be responsible
for the majority of optical spectral features of high-state CVs,
and could dominate the appearance of both UV and optical
spectra of these systems.

%% Balmer jump

% {\bf
% \noindent Notes on Section 7:

% \begin{itemize}
% 	\item

% \noindent Potential Additional Figures:

% \begin{itemize}
% 	\item 	
% }

% \subsection{Future Work}

% To further investigate the true effect of the wind
% we hope that high quality, broadband spectra of 
% Nova-like variables at a range of inclinations 
% will be taken to examine

% In addition to this project, we have started to apply the macro atom scheme to QSOs in order to build on the work of Higginbottom et al. (2013). In particular, we hope that the macro-atom scheme will enable the
% model to produce significant Lyman-$\alpha$ emission, as is observed in QSOs.

\subsection*{Acknowledgements}
The work of JHM, NH and CK is supported by the Science and Technology Facilities Council (STFC), 
via studentships and a consolidated grant, respectively. We would like to thank Juan Hernandez Santisteban and Ivan Hubeny 
for useful discussions. We acknowledge the use of the IRIDIS High Performance Computing Facility, 
and associated support services at the University of Southampton, in the completion of this work.

\newpage
\newpage
\bibliography{mybib.bib}

\end{document}