### 2015-03-25 response/supplement_response.tex
working on response

\documentclass{article}
%% declare any packages used
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{color} 
\usepackage{pdfpages}
\usepackage{appendix}
\usepackage{subfigure}
\usepackage{amsmath}
\usepackage{amssymb}

\begin{document}

\title{Supplementary material for referee\\MNRAS: MN-15-0213-MJ}
\maketitle

\begin{figure}	%fullpage
\centering
\includegraphics[width=1.0\textwidth]{pops.png}
\caption{
}
\label{novalikes}
\end{figure}

\end{document}